---
title: COVID-19 in Spain
author: Yawen Zhang
date: '2020-09-14'
slug: covid-19-in-spain
categories: []
tags: []
subtitle: ''
summary: ''
authors: []
lastmod: '2020-09-14T08:22:49+08:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
# devtools::install_github("joachim-gassen/tidycovid19")
library(tidyverse)
library(dplyr)
library(tidycovid19)
library(zoo)
library(shiny)
library(ggplot2)
library(plotly)
library(scales)
library(lubridate)
library(highcharter)
library(leaflet)
library(kableExtra)
```


```{r}
read <- read.csv("data/WHO-COVID-19-global-data.csv")
```


```{r}
# df <- download_merged_data(cached = TRUE, silent = TRUE)
# write_csv(df, path = "/Users/yawenzhang/Documents/5523 Communicating with data/my-blog-yawenzhang9701/content/post/Spain_COVID19/data/df.csv")
df <- read_csv("data/df.csv") #Save data
```

## Spain Confirmed Cases Outlier and Analysis


```{r}
our <- read.csv("data/Bing-COVID19-Data.csv") %>%
  filter(Country_Region == "Spain") %>%
  filter(AdminRegion1 %in% c("Madrid", "Catalonia", "Andalusia", "Aragon", "Basque Country", "Valencia", "Castilla Leon", "Castile-La Mancha" ,"Balearic Islands" ,"Murcia", "Galicia", "Canary Islands", "Navarre", "Extremadura", "Cantabria", "La Rioja", "Asturias", "Melilla", "Ceuta"))
```



Coronavirus crisis is one of the darkest and dramatic moments in recent Spanish history which the virus is spreading faster than in any neighboring nation. From the Map \@ref(fig:map) below, Madrid, the Spanish region with the highest number of cases, had the confirmed case of 81047 in total. After that, Catalonia became the second highest number of cases until Sep 13th. 


```{r map, fig.cap="Total Confirmed Cases in Each Region"}

post <- leaflet() %>%
            #use setView() to choose the  map center and zoom level
            setView(lat = 37.45824, lng = -4.56819,  zoom = 5) %>%
            addTiles() %>%
            addCircleMarkers(
                data = our,
                # set layer-Id to get id in click event
                layerId = ~AdminRegion1,
                lng = ~Longitude,
                lat = ~Latitude,
                radius = ~(Confirmed/3000), # to scale down
                label = ~paste("Region",AdminRegion1,",Case Number",Confirmed), #  show label name when mouse hovers
                weight = 2,
                color = "#BA5370"
            )
widgetframe::frameWidget(post, width = "100%")
```



```{r}
# our %>%
#   dplyr::select(Updated, Confirmed,AdminRegion1)%>%
#   filter(Updated == "09/13/2020") %>%
#   arrange(desc(Confirmed)) %>%
#   rename(Region = "AdminRegion1", "Total Confirmed" = "Confirmed", Date = "Updated") %>%
#   # dplyr::select(-Updated) %>%
#   filter(Region %in% c("Madrid", "Catalonia", "Andalusia", "Basque Country", "Valencia", "Castilla Leon")) %>%
#   pivot_wider(id_cols = Date, names_from = Region, values_from = "Total Confirmed") %>%
#   kable() %>%
#   kable_styling(bootstrap_options =
#                   c("striped", "condensed"),
#                   full_width = F,
#                   position = "center")
  
```

However, this set of number does not show the full picture of Spain’s coronavirus crisis. From figure \@ref(fig:line), cases for most days is zero or not available, how the coronavirus outbreak is developing through the data is not accurate. [ELPAIS](https://english.elpais.com/society/2020-04-05/spains-coronavirus-crisis-why-the-numbers-are-failing-to-show-the-full-picture.html)[@Inaccurate_cases] explain the potential reason for this situation. To be specific, the hospital may spend more than one week to process person contracting the disease, some symptoms presents and ask for a test and getting the results. In this one-week period, this person did not get isolation and may have local transmission of the coronavirus began to increase. That’s why expert believe that the number of confirmed cases may be 10 times higher than what they have been reported. 

```{r line, fig.cap="Time serious of daily new cases in each region"}
Spain_new <- our %>%
  dplyr::select(Updated, ConfirmedChange, AdminRegion1) %>%
  filter(AdminRegion1 %in% c("Madrid", "Catalonia", "Andalusia", "Aragon", "Basque Country", "Castilla Leon", "Castile-La Mancha", "Balearic Islands")) %>%
  mutate(Date = dmy(Updated)) %>%
  rename(Region = "AdminRegion1") %>%
  hchart("line", hcaes(x = Date, y = ConfirmedChange, group = Region)) 

widgetframe::frameWidget(Spain_new, width = 720)
```

What more, what caused the coronavirus in Spain to be a crisis? [Guardian](https://www.theguardian.com/world/2020/mar/26/spain-coronavirus-response-analysis)[@Guardian] gives two fundamental reasons. 
-	First:
The main reason for the quick spread through Spain may be mundane. Hugging, kissing and animated chatter from other’s face in a few inches have become a commonplace thing. 
-	Second:
Other reason was the government reacted late and clumsily and since the financial crisis, Spain has been extremely fragile. Therefore, this country lack essential equipment. It is worth to mention that, during the crisis, protective clothing for doctors and coronavirus tests are still only just being sourced and China provided assistance within its capability to help Spain [Xinhua](http://www.xinhuanet.com/english/2020-05/13/c_139051653.htm)[@Xinhua].
  
  
  
Due to coronavirus continued deterioration, the lockdown that began on 14 March enforced with police and popular pressure [Guardian](https://www.theguardian.com/world/2020/mar/26/spain-coronavirus-response-analysis)[@Guardian]. The next following section will explore lockdown restriction to citizen daily activity frequency changes.

  

## Lockdown Restriction to Citizen Daily Activity Frequency Changes




```{r}
data_wrangled <- df %>%
  filter(country == "Spain") %>%
  mutate(confirmed_lag = lag(confirmed),
         daily_confirmed = confirmed - confirmed_lag,
         death_lag = lag(deaths),
         daily_death = deaths - death_lag,
         recovered_lag = lag(recovered),
         daily_recovered = recovered - recovered_lag) %>%
  rename(residential = "gcmr_residential", workplaces = "gcmr_workplaces", retail_recreation = "gcmr_retail_recreation", grocery_pharmacy = "gcmr_grocery_pharmacy", park = "gcmr_parks", public_station = "gcmr_transit_stations")
# data_activity <- data_wrangled %>%
#   select(date, gcmr_retail_recreation, gcmr_grocery_pharmacy, gcmr_parks, gcmr_transit_stations, gcmr_workplaces, gcmr_residential) %>% 
#   pivot_longer(cols = starts_with("gcmr"), names_to = "Activity", values_to = "gcmr")
```

```{r}
# data_activity %>%
#   ggplot() +
#   geom_smooth(aes(x = date, y = gcmr, color = Activity))
```









```{r}
lockdown_date <- data.frame(Start = c("2020-03-13", "2020-03-16", "2020-03-28", "2020-04-04", "2020-04-13", "2020-04-18", "2020-05-01", "2020-05-02", "2020-05-11"), 
                            end = c("2020-03-15", "2020-03-27", "2020-04-03", "2020-04-12", "2020-04-17", "2020-04-30", "2020-05-01", "2020-05-10", "now"), 
                            lockdown_level = c("1", "2", "3", "4", "3", "4", "3", "2", "1" ))
```

Countries around the world are in uncharted territory. The emergence and spread of COVID-19 has touched every facet of society. The Spain nationwide lockdown began at March 15, as a result, daily life comes to a halt. To be specific, public gathering party are canceled, employees were sent back to home. All sizes of business are facing many uncertainties about their prospects. 
  


The following Figure \@ref(fig:work) presents the changes of different activities regarding the lockdown restrictions. Up to now, stage four restrictions happened twice since the beginning of lockdown at March 13. Stage four was an advanced phase, in which the restriction was limited. More than 23,000 people have died in Spain after contracting coronavirus, making the country one of the worst affected in the world by the illness. At this period, thousands of employee were forced to work at home during nearly three month coronavirus lockdown. Therefore, frequency that people visited workplaces has dropped about 50% compare with before. After the fourth stage lockdown restriction ended, employees allowed to work in the physical workplace and frequency that people visit workplaces increased, still lower than the average rate. Remote working caused by coronavirus has made many people aware of the convenience. Besides, Spain government are protecting the rights of remote workers, such as distance working must be a voluntary choice or employees have the right to apply compensation for any direct or indirect costs related to remote work [ELPAIS](https://english.elpais.com/economy_and_business/2020-06-26/flexible-schedules-and-covered-expenses-spain-drafts-bill-protecting-the-rights-of-remote-workers.htmla)[@ELPAIS_economics].





```{r work, fig.cap= "The frequency that people visit workplaces and residential relative to the baseline period Jan 3 - Feb 6, 2020. Since the outbreak of COVID-19, frequency that people visit residentials was increase dramatically and decreased when the fourth stage lockdown restriction finished. In contrast, the frequency that people visit workplaces decrease significantly until the fourth stage lockdown restriction finished."}
Activity <- data_wrangled %>%
  dplyr::select(date, residential, workplaces) %>%
  pivot_longer(cols = c(residential, workplaces), names_to = "Daily_Activity", values_to = "Changes") %>%
  ggplot() +
  geom_smooth(aes(x = date, y = Changes, colour = Daily_Activity), 
              se = FALSE, 
              method = 'loess') +
  geom_rect(data = lockdown_date %>% 
              filter(lockdown_level == "4"), 
            aes(xmin = as.Date(Start), xmax = as.Date(end), ymin = -Inf, ymax = Inf), 
            alpha = 0.25, 
            fill = "#BA5370")+
  theme_light() +
  scale_x_date(breaks=date_breaks("1 month"),
               labels=date_format('%b')) +
  annotate(geom = "text", 
           x=as.Date('2020-03-20'), 
           y=0, 
           label="First Stage 4", 
           size = 4, 
           color = "red") +
  annotate(geom = "text", 
           x=as.Date('2020-05-10'), 
           y=-30, 
           label="Second Stage 4", 
           size = 4, 
           color = "red")
Activity
  # geom_rect(data = lockdown_date %>% filter(lockdown_level == "3"), aes(xmin = as.Date(Start), xmax = as.Date(end), ymin = -Inf, ymax = Inf),alpha = 0.25, fill = "#BA5370")
```





Within the lockdown period, parks remain closed. Therefore, runners and cyclists take the broad avenue as a racetrack. Most of them transfer the outdoor sports to exercise at home. In Figure \@ref(fig:daily), people rarely visit parks during the lockdown period. At the end of April, it has seen a steadily declining number of deaths per day, leading to the country's plan to broaden lockdown restrictions. Spain has released an exit strategy from the lockdown restriction in a bid to return the normality, while keeping COVID-19 under control [euronews](https://www.euronews.com/2020/04/29/coronavirus-spain-reveals-four-stage-plan-to-de-escalate-from-covid-19-lockdown)[@euronews]. With the gradual relaxation of the lockdown restriction, retail and recreation industry allowed to re-open to the public, such as sidewalk café reopened and Spaniards can exercise outdoors [CNN](https://www.cnn.com/2020/05/02/europe/spain-lockdown-coronavirus-exercise-intl/index.html)[@CNN]. As a consequence, frequency that people visit parks increase over than 50%. 




```{r daily, fig.cap="The frequency changes that people visit pharmacy, parks, retail and recreation relative  to the baseline period Jan 3 - Feb 6, 2020"}
Daily <- data_wrangled %>%
  dplyr::select(date, retail_recreation, grocery_pharmacy, park) %>%
  pivot_longer(cols = c(retail_recreation, grocery_pharmacy, park), names_to = "Activity", values_to = "Changes") %>%
  ggplot() +
  geom_smooth(aes(x = date, y = Changes, colour = Activity), se = FALSE, method = 'loess') +
  geom_rect(data = lockdown_date %>% filter(lockdown_level == "4"), aes(xmin = as.Date(Start), xmax = as.Date(end), ymin = -Inf, ymax = Inf),alpha = 0.25, fill = "#BA5370") +
  theme_light() +
  scale_x_date(breaks=date_breaks("1 month"),
               labels=date_format('%b')) +
  annotate(geom = "text", 
           x=as.Date('2020-03-20'), 
           y=50, 
           label="First Stage 4", 
           size = 4, 
           color = "red") +
  annotate(geom = "text", 
           x=as.Date('2020-05-15'), 
           y=40, 
           label="Second Stage 4", 
           size = 4, 
           color = "red")
Daily
```












# References
