---
title: Situation in Spain `r emo::ji("es")` with COVID-19
author: Yawen Zhang
date: '2020-09-14'
slug: covid-19-in-spain
categories: []
tags: []
subtitle: ''
summary: ''
authors: [Yawen Zhang]
lastmod: '2020-09-14T08:22:49+08:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
# devtools::install_github("joachim-gassen/tidycovid19")
library(tidyverse)
library(dplyr)
library(tidycovid19)
library(zoo)
library(ggplot2)
library(plotly)
library(scales)
library(lubridate)
library(highcharter)
library(leaflet)
library(kableExtra)
```


```{r}
read <- read.csv("data/WHO-COVID-19-global-data.csv")
```


```{r}
# df <- download_merged_data(cached = TRUE, silent = TRUE)
# write_csv(df, path = "/Users/yawenzhang/Documents/5523 Communicating with data/my-blog-yawenzhang9701/content/post/Spain_COVID19/data/df.csv")
df <- read_csv("data/df.csv") #Save data
```

## Coronavirus History in Spain `r emo::ji("es")`

The COVID-19 pandemic in Spain is part of the pandemic of Coronavirus Disease 2019 (**COVID‑19**). Spain has the second highest total confirmed cases in Europe after Russia. The first confirmed cases came from Germany and spread to Spain on 31 Jan 2019. Up to 15 March 2020, all 50 provinces have confirmed cases of the country. Spain became one of the most draconian countries in the world and a lockdown restriction imposed on 14 March 2020. Until 13 Sep, there are about 566,326 confirmed cases and 29,747 deaths [WHO](https://covid19.who.int/region/euro/country/es)[@WHO]. 


## Data Source and Descriptions `r emo::ji("rotating_light")`
The data is taken from the [tidycovid19](https://github.com/joachim-gassen/tidycovid19)[@Spain] package `r emo::ji("package")` created by Joachim Gassen. The `tidycovid19` package was created to provide transparent access to various data sources. Then, used `download_merged_data` to merge the dataset. This merge dataset contains data of coronavirus confirmed, recovered, death original collected from [Johns Hopkins University](https://github.com/CSSEGISandData/COVID-19)[@Johns_Hopkins].Some other variables contains the frequency places are visited comes from [Google Commmunity Mobility Reports](https://www.google.com/covid19/mobility/)[@Google]. The second set of data comes from [Bing](https://github.com/microsoft/Bing-COVID-19-Data)[@Bing] which contains all the regions location and cases in Spain. The raw data was stored in two CSV files separately. The data collection date for this blog until September 13, 2020. The following map and plots present these variables in the data set and also contain descriptions.



## Spain Confirmed Cases and Analysis `r emo::ji("world_map")`


```{r}
our <- read.csv("data/Bing-COVID19-Data.csv") %>%
  filter(Country_Region == "Spain") %>%
  filter(AdminRegion1 %in% c("Madrid", "Catalonia", "Andalusia", "Aragon", "Basque Country", "Valencia", "Castilla Leon", "Castile-La Mancha" ,"Balearic Islands" ,"Murcia", "Galicia", "Canary Islands", "Navarre", "Extremadura", "Cantabria", "La Rioja", "Asturias", "Melilla", "Ceuta"))
```



Coronavirus crisis is one of the darkest and dramatic moments in recent Spanish history which the virus is spreading faster than any neighboring nation. From the Map \@ref(fig:map) below, Madrid, the Spanish region with the highest number of cases, had the confirmed case of 81,047 in total. After that, Catalonia became the second highest number of cases up to 62,524 until September 13th.  


```{r map, fig.cap="Total Confirmed Cases in Each Region"}

post <- leaflet() %>%
            #use setView() to choose the  map center and zoom level
            setView(lat = 41.79675, lng = 1.53172,  zoom = 6) %>%
            addTiles() %>%
            addCircleMarkers(
                data = our,
                # set layer-Id to get id in click event
                layerId = ~AdminRegion1,
                lng = ~Longitude,
                lat = ~Latitude,
                radius = ~(Confirmed/1500), # to scale down
                label = ~paste("Region",AdminRegion1,",Case Number",Confirmed), #  show label name when mouse hovers
                weight = 2,
                color = "red"
            )
widgetframe::frameWidget(post, width = "160%")
```



```{r}
# our %>%
#   dplyr::select(Updated, Confirmed,AdminRegion1)%>%
#   filter(Updated == "09/13/2020") %>%
#   arrange(desc(Confirmed)) %>%
#   rename(Region = "AdminRegion1", "Total Confirmed" = "Confirmed", Date = "Updated") %>%
#   # dplyr::select(-Updated) %>%
#   filter(Region %in% c("Madrid", "Catalonia", "Andalusia", "Basque Country", "Valencia", "Castilla Leon")) %>%
#   pivot_wider(id_cols = Date, names_from = Region, values_from = "Total Confirmed") %>%
#   kable() %>%
#   kable_styling(bootstrap_options =
#                   c("striped", "condensed"),
#                   full_width = F,
#                   position = "center")
  
```

However, this set of number does not show the full picture of Spain’s coronavirus crisis. Figure \@ref(fig:line) does not present daily new cases for all regions, otherwise the plot will be extremely complex and hard to read. Cases for most days is zero or not available. Therefore, how the coronavirus outbreak is developing through the data is not accurate. [ELPAIS](https://english.elpais.com/society/2020-04-05/spains-coronavirus-crisis-why-the-numbers-are-failing-to-show-the-full-picture.html)[@Inaccurate_cases] explain the potential reason for this situation. To be specific, the hospital `r emo::ji("hospital")` may spend more than one week to process person contracting the disease, some symptoms presents and ask for a test and getting the results. In this one-week period, this person did not get isolation and may have local transmission of the coronavirus began to increase. That’s why experts believe that the number of confirmed cases may be 10 times higher than what they have been reported. 

```{r line, fig.cap="Time serious of daily new cases in each region"}
Spain_new <- our %>%
  dplyr::select(Updated, ConfirmedChange, AdminRegion1) %>%
  filter(AdminRegion1 %in% c("Madrid", "Catalonia", "Andalusia", "Aragon", "Basque Country", "Castilla Leon", "Castile-La Mancha", "Balearic Islands")) %>%
  mutate(Date = dmy(Updated)) %>%
  rename(Region = "AdminRegion1") %>%
  hchart("line", hcaes(x = Date, y = ConfirmedChange, group = Region)) %>%
  hc_title(text = "Time serious of daily new cases in each region",
           margin = 20,
           style = list(color = "black", useHTML = TRUE))

widgetframe::frameWidget(Spain_new, width = 720)
```

What more, what caused the coronavirus in Spain to be a crisis? [Guardian](https://www.theguardian.com/world/2020/mar/26/spain-coronavirus-response-analysis)[@Guardian] gives two fundamental reasons. 
  
* Mundane:
The main reason for the quick spread through Spain may be mundane. Hugging, kissing and animated chatter from other’s face in a few inches have become a commonplace thing. 
* Government:
Other reason was the government reacted late and clumsily. 


Since the financial crisis, Spain has been extremely fragile and this country lack essential equipment. During the crisis, protective clothing for doctors and coronavirus tests were still only just being sourced and China provided assistance within its capability to help Spain (http://www.xinhuanet.com/english/2020-05/13/c_139051653.htm)[@Xinhua].
  
  
  
Due to coronavirus continued deterioration, the lockdown that began on 14 March enforced with police and popular pressure [Guardian](https://www.theguardian.com/world/2020/mar/26/spain-coronavirus-response-analysis)[@Guardian]. The next following section will explore lockdown restriction to citizen daily activity frequency changes.

  

## Lockdown Restriction to Citizen Daily Activity Frequency Changes `r emo::ji("footprints")`




```{r}
data_wrangled <- df %>%
  filter(country == "Spain") %>%
  mutate(confirmed_lag = lag(confirmed),
         daily_confirmed = confirmed - confirmed_lag,
         death_lag = lag(deaths),
         daily_death = deaths - death_lag,
         recovered_lag = lag(recovered),
         daily_recovered = recovered - recovered_lag) %>%
  rename(residential = "gcmr_residential", workplaces = "gcmr_workplaces", retail_recreation = "gcmr_retail_recreation", grocery_pharmacy = "gcmr_grocery_pharmacy", park = "gcmr_parks", public_station = "gcmr_transit_stations")
# data_activity <- data_wrangled %>%
#   select(date, gcmr_retail_recreation, gcmr_grocery_pharmacy, gcmr_parks, gcmr_transit_stations, gcmr_workplaces, gcmr_residential) %>% 
#   pivot_longer(cols = starts_with("gcmr"), names_to = "Activity", values_to = "gcmr")
```

```{r}
# data_activity %>%
#   ggplot() +
#   geom_smooth(aes(x = date, y = gcmr, color = Activity))
```









```{r}
lockdown_date <- data.frame(Start = c("2020-03-13", "2020-03-16", "2020-03-28", "2020-04-04", "2020-04-13", "2020-04-18", "2020-05-01", "2020-05-02", "2020-05-11"), 
                            end = c("2020-03-15", "2020-03-27", "2020-04-03", "2020-04-12", "2020-04-17", "2020-04-30", "2020-05-01", "2020-05-10", "now"), 
                            lockdown_level = c("1", "2", "3", "4", "3", "4", "3", "2", "1" ))
```

Countries around the world are in uncharted territory. The emergence and spread of COVID-19 has touched every facet of society. The Spain nationwide lockdown began at March 14, as a result, daily life comes to a halt, such as public gathering party are canceled, employees were sent back to home. All sizes of business are facing many uncertainties about their prospects. 
  


The following Figure \@ref(fig:work) and Figure \@ref(fig:daily) presents the changes of different activities regarding the lockdown restrictions`r emo::ji("lock")`. Up to now, stage four restrictions happened twice since the beginning of lockdown at March 14. Stage four was an advanced phase, in which the restriction was limited. More than 23,000 people have died in Spain after contracting coronavirus, making the country one of the worst affected in the world by the illness. At this period, thousands of employee were forced to work at home during nearly three-month coronavirus lockdown. Therefore, frequency that people visited workplaces has dropped about 50% compare with before. After the fourth stage lockdown restriction ended, employees allowed to work in the physical workplace and frequency that people visit workplaces increased, but still lower than the average rate. Remote working caused by coronavirus has made many people aware of the convenience. Besides, Spain government are protecting the rights of remote workers, such as distance working must be a voluntary choice or employees have the right to apply compensation for any direct or indirect costs related to remote work [ELPAIS](https://english.elpais.com/economy_and_business/2020-06-26/flexible-schedules-and-covered-expenses-spain-drafts-bill-protecting-the-rights-of-remote-workers.htmla)[@ELPAIS_economics].





```{r work, fig.cap= "The frequency that people visit workplaces and residential relative to the baseline period Jan 3 - Feb 6, 2020."}
Activity <- data_wrangled %>%
  dplyr::select(date, residential, workplaces) %>%
  pivot_longer(cols = c(residential, workplaces), names_to = "Daily_Activity", values_to = "Changes") %>%
  ggplot() +
  geom_smooth(aes(x = date, y = Changes, colour = Daily_Activity), 
              se = FALSE, 
              method = 'loess') +
  geom_rect(data = lockdown_date %>% 
              filter(lockdown_level == "4"), 
            aes(xmin = as.Date(Start), xmax = as.Date(end), ymin = -Inf, ymax = Inf), 
            alpha = 0.25, 
            fill = "#BA5370")+
  theme_light() +
  scale_x_date(breaks=date_breaks("1 month"),
               labels=date_format('%b')) +
  annotate(geom = "text", 
           x=as.Date('2020-03-20'), 
           y=0, 
           label="First Stage 4", 
           size = 4, 
           color = "red") +
  annotate(geom = "text", 
           x=as.Date('2020-05-10'), 
           y=-30, 
           label="Second Stage 4", 
           size = 4, 
           color = "red") +
  ggtitle("The frequency that people visit workplaces and residential") +
  theme(plot.title = element_text(size = 14, hjust = 0.5))
Activity
  # geom_rect(data = lockdown_date %>% filter(lockdown_level == "3"), aes(xmin = as.Date(Start), xmax = as.Date(end), ymin = -Inf, ymax = Inf),alpha = 0.25, fill = "#BA5370")
```





Within the lockdown period, parks closed. Most of them transfer the outdoor sports to exercise at home. In Figure \@ref(fig:daily), people rarely visit parks during the lockdown period. At the end of April, it has seen a steadily declining number of cases per day, leading to the country's plan to broaden lockdown restrictions. Spain has released an exit strategy from the lockdown restriction in a bid to return the normality, while keeping COVID-19 under control [euronews](https://www.euronews.com/2020/04/29/coronavirus-spain-reveals-four-stage-plan-to-de-escalate-from-covid-19-lockdown)[@euronews]. With the gradual relaxation of the lockdown restriction, retail and recreation industry allowed to re-open to the public, such as sidewalk café reopened and Spaniards can exercise outdoors [CNN](https://www.cnn.com/2020/05/02/europe/spain-lockdown-coronavirus-exercise-intl/index.html)[@CNN]. As a consequence, frequency that people visit parks increase over than 50% which is higher than any other activities. 




```{r daily, fig.cap="The frequency changes that people visit pharmacy, parks, retail and recreation relative  to the baseline period Jan 3 - Feb 6, 2020"}
Daily <- data_wrangled %>%
  dplyr::select(date, retail_recreation, grocery_pharmacy, park) %>%
  pivot_longer(cols = c(retail_recreation, grocery_pharmacy, park), names_to = "Activity", values_to = "Changes") %>%
  ggplot() +
  geom_smooth(aes(x = date, y = Changes, colour = Activity), se = FALSE, method = 'loess') +
  geom_rect(data = lockdown_date %>% filter(lockdown_level == "4"), aes(xmin = as.Date(Start), xmax = as.Date(end), ymin = -Inf, ymax = Inf),alpha = 0.25, fill = "#BA5370") +
  theme_light() +
  scale_x_date(breaks=date_breaks("1 month"),
               labels=date_format('%b')) +
  annotate(geom = "text", 
           x=as.Date('2020-03-20'), 
           y=50, 
           label="First Stage 4", 
           size = 4, 
           color = "red") +
  annotate(geom = "text", 
           x=as.Date('2020-05-15'), 
           y=40, 
           label="Second Stage 4", 
           size = 4, 
           color = "red") +
  ggtitle("The frequency changes that people visit pharmacy, parks, retail and recreation")+
  theme(plot.title = element_text(size = 11, hjust = 0.5))
Daily
```

For now, Spain has decreased to the first level of lockdown, daily life gradually return to normal. However, face mask are now obligatory at ALL times and keep two meters social distance. Some regions such as Aragon Madrid and Cataluña are currently experiencing outbreaks that has brought back lockdown in some towns and regions. The government is also making regulations based on the current situation. 


## Acknowledgements
The packages used for producing this post are:

tidyverse [@tidyverse], knitr [@knitr], kableExtra [@kableExtra], dplyr [@dplyr], tidycovid19[@tidycovid19], zoo[@zoo], ggplot[@ggplot],scales[@scales], highcharter[@highcharter], leaflet[@leaflet], lubridate[@lubridate]. 

The raw data comes from two github repositories: tidycovid19[@Spain] and Bing[@Bing].





## References
